---
title       : Vectors
description : The most basic unit to store data in R are vectors. We'll learn about creating and working with them in this chapter.

---
## Numeric Vectors

```yaml
type: NormalExercise
key: eb99436e3b
lang: r
xp: 100
skills: 1
```
A vector is a series of values, all of the same type. They are the most basic data type in R and can hold numeric data, character data, or logical data. In R, you can create a vector with the combine function `c()`. You place the vector elements separated by a comma between the parentheses. For example a numeric vector would look something like this: 

```{r}
cost <- c(50, 75, 90, 100, 150)
```

`@instructions`
Use the function `c()` to create a numeric vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto.
The average high temperatures are `35, 88, 42, 84, 81, and 30` degrees Fahrenheit. Call the object `temp`.

`@hint`
Make sure your numbers are within parenthesis and have commas separating them. 

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Here is an example creating a numeric vector named cost
cost <- c(50, 75, 90, 100, 150)

# Create a numeric vector to store the temperatures listed in the instructions into a vector named temp
# Make sure to follow the same order in the instructions

```

`@solution`
```{r}
# Here is an example creating a numeric vector named cost
cost <- c(50, 75, 90, 100, 150)

# Create a numeric vector to store the temperatures listed in the instructions into a vector named temp
# Make sure to follow the same order in the instructions
temp <- c(35, 88, 42, 84, 81, 30)
```

`@sct`
```{r}
test_error()
test_object("temp", incorrect_msg = "Make sure the numbers you entered match the numbers in the instructions.")
success_msg("Good job! Do you want to now try creating a character vector")
```

---
## Character vectors

As in the previous question, we are going to create a vector. Only this time, we learn to create *character* vectors. The main difference is that these have to be written as strings and so the names are enclosed within double quotes. 

A *character* vector would look something like this:

```{r}
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
```

```yaml
type: NormalExercise
key: a99d0bbace
lang: r
xp: 100
skills: 1
```


`@instructions`
The temperatures we stored in `temp` are from the cities Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto.
- Create a vector with these city names and call the object `city`. 

`@hint`
Be sure to place the names within double quotes. 

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# here is an example of how to create a character vector
food <- c("pizza", "burgers", "salads", "cheese", "pasta")

# Create a character vector called city to store the city names
# Make sure to follow the same order as in the instructions

```

`@solution`
```{r}
# here is an example of how to create a character vector
food <- c("pizza", "burgers", "salads", "cheese", "pasta")

# Create a character vector called city to store the city names
# Make sure to follow the same order as in the instructions
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

```

`@sct`
```{r}
test_error()
test_object("city", incorrect_msg = "Make sure to use the correct order.",undefined_msg = "Each of the city names need double quotes.")
success_msg("Awesome! Now you've learnt to store both numeric and character vectors.")

```

---
## Connecting Numeric and Character Vectors 

We have successfully assigned the temperatures as *numeric* values to `temp` and the city names as *character* values to `city`. But can we associate the temperature to it's related city? Yes! We can do so using a code we already know - `names`. We assign names to the *numeric* values.  

It would look like this:

```{r}
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food
```

```yaml
type: NormalExercise
key: 62bba83b36
lang: r
xp: 100
skills: 1
```


`@instructions`
Use the `names` function and the objects defined in the previous exercises to associate the temperature data with its corresponding city.
(You can go back to the previous questions and copy the objects stored.) Note: to see what happened, after assigning the city
names to the temp vector, try printing the temp vector to understand how the names are associated with elements of temp.

`@hint`
Assign names to the numeric values. 

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# Associate the cost values with its corresponding food item
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food

# You already wrote this code
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Associate the temperature values with its corresponding city

```

`@solution`
```{r}
# Associate the cost values with its corresponding food item
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food

# You already wrote this code
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Associate the temperature values with its corresponding city
names(temp) <- city
```

`@sct`
```{r}
test_error()
test_object("temp", incorrect_msg = "Check to match the numbers from the question!")
test_function("names",index=2,incorrect_msg = "The numeric vector should be in the parenthesis. Also, don't delete the sample code.") 
test_student_typed("names(temp)<-city", not_typed_msg = "We want you to assign the names stored in `city` using the names function. And remember to use `<-` rather than `=`.")
success_msg("Great job! We now know the temperatures of the respective cities.")
```

---
## Subsetting vectors

If we want to display only selected values from the object, R can help us do that easily. 

For example, if we want to see the cost of the last 3 items in our food list, we would type:

```{r}
cost[3:5]
```

Note here, that we could also type `cost[c(3,4,5)]` and get the same result. 
The `:` operator helps us condense the code and get consecutive values. 

```yaml
type: NormalExercise
key: b9d09364f0
lang: r
xp: 100
skills: 1
```


`@instructions`
We will learn to subset using several special operators.
- Use the `[` and `:` operators to access the temperature of the first three cities in the list, which are already stored in `temp`.

`@hint`
Square brackets (`[`) and `:` are the important bits in this.

`@pre_exercise_code`
```{r}
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food

temp <- c(35, 88, 42, 84, 81, 30) 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city

```

`@sample_code`
```{r}
# cost of the last 3 items in our food list:
cost[3:5]

# temperatures of the first three cities in the list:

```

`@solution`
```{r}
# cost of the last 3 items in our food list:
cost[3:5]

# temperatures of the first three cities in the list:
temp[1:3]

```

`@sct`
```{r}
test_error()
test_output_contains("temp[1:3]", incorrect_msg = "The square brackets and colon operators are essential to this command.")
test_student_typed(c("temp[1:3]", "temp[c(1,2,3)]", "temp[seq(1,3)]"), not_typed_msg = "Remember to use the brackets `[` and `]`. And inside the brackets you should have a vector with entries 1, 2 and 3 which you can generate with `:`, `c` or `seq`")
success_msg("Awesome! You've learnt how to make your work easier. But make sure to keep it simple. ")
```

---
## Subsetting vectors continued...

In the previous question, we accessed the temperature for consecutive cities (1st three). But what if we want to access the temperatures for any 2 specific cities?

An example:
To access the cost of `pizza` (1st) and `pasta` (5th food item) in our list, the code would be:
 
```{r}
cost[c(1,5)]
```

```yaml
type: NormalExercise
key: 16e3329348
lang: r
xp: 100
skills: 1
```


`@instructions`
This time we will access our object using just the `[` operator.
Use the `[` operator to access the temperature of Paris and San Juan already stored in the `temp` object.

`@hint`
Don't forget about the combine, `c()`, function.

`@pre_exercise_code`
```{r}
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food

temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city
```

`@sample_code`
```{r}
# Access the cost of pizza and pasta from our food list 
cost[c(1,5)]

# Define temp
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city

# Access the temperatures of Paris and San Juan

```

`@solution`
```{r}
# Access the cost of pizza and pasta from our food list 
cost[c(1,5)]

# Define temp
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city

# Access the temperatures of Paris and San Juan
temp[c(3,5)]

```

`@sct`
```{r}
test_error()
test_output_contains("temp[c(3,5)]", incorrect_msg = "You need to use `c` and parenthesis inside the square brackets. Look at how `temp` is defined in the sample code to determine the entries associated with Paris and San Juan")
success_msg("Isn't that awesome! Let's move to the next exercise!")
```

---
## Sequences 

The `:` operator helps us create sequences of numbers. 
For example, `32:99` would create a list of numbers from 32 to 99. 

Then, if we want to know the length of this sequence, all we need to do use the `length` command. 


```yaml
type: NormalExercise
key: 90eed94daa
lang: r
xp: 100
skills: 1
```


`@instructions`
This time we will use just the `:` operator.
Use the `:` operator to create a sequence of consecutive integers starting 12 and ending at 73 and save it in an object `x`.

`@hint`
Save the sequence in an object called `x`. 

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# Create a vector m of integers that starts at 32 and ends at 99.
m <- 32:99

# Determine the length of object m.
length(m)

# Create a vector x of integers that starts 12 and ends at 73.

# Determine the length of object x.


```

`@solution`
```{r}
# Create a vector m of integers that starts at 32 and ends at 99.
m <- 32:99

# Determine the length of object m.
length(m)

# Create a vector x of integers that starts 12 and ends at 73.
x <- 12:73

# Determine the length of object x.
length(x)

```

`@sct`
```{r}
test_error()
test_object("x", undefined_msg = "Make sure to define `x`!", incorrect_msg = "Have you correctly assigned 12:73 to `x`?")
test_function("length",index=2,not_called_msg="Make sure to call length on x and don't delete the sample code")
test_output_contains("length(x)", incorrect_msg = "Make sure you are returning the correct length.")
success_msg("Good job! Let's move to the next question!")
```

---
## Sequences continued...

We can also create different types of sequences in R. 
For example, in `seq(7, 49, 7)`, the first argument defines the start, and the second the end. The default is to go up in increments of 1, but a third argument lets us tell it by what interval.


```yaml
type: NormalExercise
key: c6535a161e
lang: r
xp: 100
skills: 1
```


`@instructions`
We will learn how to use the `seq()` function in this question.
- Create a vector containing all the positive odd numbers smaller than 100. The numbers should be in ascendint order.
`@hint`
If you start at 1 and jump by 2, you`ll get the positive odd numbers. 

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Create a vector with the multiples of 7, smaller than 50.
seq(7, 49, 7) 

# Create a vector containing all the positive odd numbers smaller than 100.
# The numbers should be in ascending order

```

`@solution`
```{r}
# Create a vector with the multiples of 7, smaller than 50. 
seq(7, 49, 7) 

# Create a vector containing all the positive odd numbers smaller than 100.
# The numbers should be in ascending order
seq(1, 100, 2)

```

`@sct`
```{r}
test_error()
test_output_contains("seq(1,100,2)", incorrect_msg = "Make sure the third argument has the value the sequence needs to jump by.")
success_msg("Awesome! Let us 'jump' to the next exercise.")
```

---
## Sequences and length 

The second argument of the function `seq` is actually a maximum, not necessarily the end. So if we type

```{r}
seq(7, 50, 7)
```

we actually get the same vector of integers as if we type

```{r}
seq(7, 49, 7)
```

This can be useful because sometimes all we want are sequential numbers  that are smaller than some value. 

Let's look at an example.

```yaml
type: NormalExercise
key: c8deb909eb
lang: r
xp: 100
skills: 1
```


`@instructions`
Create a vector of numbers that starts at 6, does not go beyond 55, and adds numbers in increments of 4/7.
So the first three numbers will be 6, 6+4/7, and 6+8/7. How many numbers does the list have? Use only one line of code to answer both questions.

`@hint`
Use `length` and then `seq` within parentheses.

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# We can a vector with the multiples of 7, smaller than 50 like this 
seq(7, 49, 7) 

# But note that the second argument does not need to be last number.
# It simply determines the maximum value permitted.
# so the following line of code produces the same vector as seq(7, 49, 7)
seq(7, 50, 7)

# Create a sequence of numbers from 6 to 55, with 4/7 increments and determine its length


```

`@solution`
```{r}
# We can a vector with the multiples of 7, smaller than 50 like this 
seq(7, 49, 7) 

# But note that the second argument does not need to be last number.
# It simply determines the maximum value permitted.
# so the following line of code produces the same vector as seq(7, 49, 7)
seq(7, 50, 7)

# Create a sequence of numbers from 6 to 55, with 4/7 increments and determine its length
length(seq(6, 55, 4/7))

```

`@sct`
```{r}
test_error()
test_output_contains("length(seq(6, 55, 4/7))", incorrect_msg = "Check the code again.")
success_msg("Great job! Looks like you're getting good with sequences!")
```

---
## Sequences of certain length

The `seq()` function has another useful argument. The argument *length.out*. This argument lets us generate sequences that are increasing by the same amount but are of the prespecified length. 

For example, this line of code 

```{r}
x <- seq(0, 100, length.out = 5)
```

produces the numbers 0, 25, 50, 75, 100. 

Let's create a vector and see what is the class of the object produced. 


```yaml
type: NormalExercise
key: a005f7b688
lang: r
xp: 100
skills: 1
```


`@instructions`
- Determine the class of a vector generated with `seq` using the *length.out* argument. 
- Specifically, what is the class of the following object `a <- seq(1, 10, length.out = 100)`?

`@hint`
Create the sequence, store it in an object, then check the class.

`@pre_exercise_code`
```{r}
#no pec 
```

`@sample_code`
```{r}
# Store the sequence in the object a


# Determine the class of a


```

`@solution`
```{r}
# Store the sequence in the object a
a <- seq(1, 10, length.out = 100)

# Determine the class of a
class(a)

```

`@sct`
```{r}
test_error()
test_object("a", undefined_msg = "Make sure you store the sequence in object `a`.", incorrect_msg = "There's something wrong with the sequence.")
test_function("class")
success_msg("Good job! Let us move on to the next exercise!")
```

---
## Integers

We have discussed the *numeric* class. We just saw that the `seq` function can generate objects of this class. For another example type 

```{r}
class(seq(1, 10, 0.5))
```
into the console and note that the `class` is *numeric*. R has another type of vector we have not described, the *integer* class. You can create an *integer* by adding the letter `L` after a whole number. If you type

```{r}
class(3L)
```

in the console, you see this is an *integer* and not a *numeric*. For most  practical purposes, integers and numerics are indistinguishable. For example 3, the integer, minus 3 the numeric is 0. To see this type this in the console

```{r}
3L - 3
```

The main difference is that integers occupy less space in the computer memory, so for big computations using integers can have a substantial impact.


```yaml
type: NormalExercise
key: b5c6d6d5ab
lang: r
xp: 100
skills: 1
```


`@instructions`
Does this change depending on what we store in an object?
What is the class of the following object `a <- seq(1, 10)`? 

`@hint`
Note that there are only integer values within the seq parenthesis. 

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# Store the sequence in the object a


# Determine the class of a

```

`@solution`
```{r}
# Store the sequence in the object a
a <- seq(1, 10)

# Determine the class of a
class(a)

```

`@sct`
```{r}
test_error()
test_object("a", undefined_msg = "Make sure you store the sequence in object `a`.", incorrect_msg = "There's something wrong with the sequence.")
test_function("class")
test_output_contains("class(a)", incorrect_msg = "Make sure you call the `class` function on the correct object.")
success_msg("Good job! Let's move on to the next exercise!")
```

---
## Integers and Numerics

Let's confirm that `1L` is an *integer* not a *numeric*.

```yaml
type: NormalExercise
key: 8517a900a9
lang: r
xp: 100
skills: 1
```


`@instructions`
Confirm that the `class` of `1` is *numeric* and the `class` of `1L` is *integer*.

`@hint`
Just call the function `class` with `1L` as an argument.

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# Check the class of 1, assigned to the object a
class(1)

# Confirm the class of 1L is integer

```

`@solution`
```{r}
# Check the class of 1, assigned to the object a
class(1)

# Confirm the class of 1L is integer
class(1L)

```

`@sct`
```{r}
test_error()
test_output_contains("class(1L)", incorrect_msg = "Check the class of `1L` to make sure the L makes it an integer.")
success_msg("Great, now that you know how to force a number to become an integer, let us move on to one more fun thing in this module.")
```

---
## Coercion

The concept of coercion is a very important one. Watching the video, we learned that when an entry does not match what an R function is expecting, R tries to guess what we meant before throwing an error. This might get confusing at times. 

As we've discussed in earlier questions, there are numeric and character vectors. The character vectors are placed in quotes and the numerics are not. 

We can avoid issues with coercion in R by changing characters to numerics and vice-versa. This is known as typecasting. The code, `as.numeric(x)` helps us convert character strings to integers. There is an equivalent function that converts integers to strings, `as.character(x)`.

Let's practice doing this!

```yaml
type: NormalExercise
key: 6508787042
lang: r
xp: 100
skills: 1
```


`@instructions`
- Define the following vector:

```{r}
x <- c(1, 3, 5,"a")
```

- Notice that `x` has been coerced into a character string.
- Convert `x` to a vector of integers using the `as.numeric()` function.

`@hint`
Make sure you put the numbers within "". 

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# Define the vector x
x <- c(1, 3, 5,"a")

# Note that the x is character vector
x

# Typecast the vector to get an integer vector
# You will get a warning but that is ok

```

`@solution`
```{r}
# Define the vector x
x <- c(1, 3, 5,"a")

# Note that the x is character vector
x

# Typecast the vector to get an integer vector
# You will get a warning but that is ok
x = as.numeric(x)

```

`@sct`
```{r}
test_error()
test_object("x", undefined_msg = "Make sure to define x first.", incorrect_msg = "Assign the vector back to itself in `x`.")
test_function("as.numeric")
success_msg("Awesome! Doesn't that feel wonderful? Now, you can try changing it back to character, using the code `as.character(x)`.")
```

---
## End of Assessment 3

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: d596636416
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/2a638f36b59f4897b1bd6867ad58f63e/25784d8d534443cdacc07cb670376dff/?activate_block_id=block-v1%3AHarvardX%2BPH125.1x%2B2T2017%2Btype%40sequential%2Bblock%4025784d8d534443cdacc07cb670376dff'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K. WARNING: if you continue the assessments by clicking on the arrow to get the next exercise or by hitting Ctrl-K your assessments may NOT get scored.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```


```
