---
title       : Basic Plots
description : We engage in the basics of plot-making in R, with scatterplots, histograms and box plots.
---
## Scatterplots

We made a plot of total murders versus population and noted a strong relationship: not surprisingly states with larger populations had more murders. You can run the code in the console to get the plot.

```{r}
library(dslabs)
data(murders)

population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total

plot(population_in_millions, total_gun_murders)
```

Note that many states have populations below 5 million and are bunched up in the plot. We may gain further insights from making this plot in the log scale. 

```yaml
type: NormalExercise
key: c535f6f10b
lang: r
xp: 100
skills: 1
```

`@instructions`
- Transform the variables using the log, to the base 10, transformation
- Plot the log transformed total murders versus population

`@hint`
Once you define the objects `log10_population` and `log10_total_gun_murders` you plot them against each other using
`plot(log10_population, log10_total_gun_murders)`

```

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Load the datasets and define some variables
library(dslabs)
data(murders)

population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total

plot(population_in_millions, total_gun_murders)

# Transform population using the log10 transformation and save to object log10_population

# Transform total gun murders using log10 transformation and save to object log10_total_gun_murders

# Create a scatterplot with the log scale transformed population and murders 

```

`@solution`
```{r}
# Load the datasets and define some variables
library(dslabs)
data(murders)
population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total
plot(population_in_millions, total_gun_murders)

# Transform population using the log10 transformation and save to object log10_population
log10_population<- log10(murders$population)

# Transform total gun murders using log10 transformation and save to object log10_total_gun_murders
log10_total_gun_murders <- log10(murders$total)

# Create a scatterplot with the log scale transformed population and murders 
plot(log10_population, log10_total_gun_murders)

```

`@sct`
```{r}
test_error()
test_object("log10_population", undefined_msg = "Define log10_population first.", incorrect_msg = "Transform variable using log10. Since we are taking the log do not divide by a million.")
test_object("log10_total_gun_murders", undefined_msg = "Define log10_total_gun_murders first.", incorrect_msg = "Transform variable using log10.")
test_function("plot",args=c("x","y"), incorrect_msg = "Scatter population on murders.")
success_msg("Doesn't that plot look better!")
```

---
## Histograms

Now we are going to make a histogram.

```yaml
type: NormalExercise
key: 6d782c33a3
lang: r
xp: 100
skills: 1
```


`@instructions`
- Compute the population in millions and save it to the object `population_in_millions`
- Create a histogram of the state populations using the function `hist`

`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Store the population in millions and save to population_in_millions 
population_in_millions <- murders$population/10^6


# Create a histogram of this variable

```

`@solution`
```{r}
# Store the population in millions (same as previous question)
population_in_millions <- murders$population/10^6

# Create a histogram of this variable
hist(population_in_millions)

```

`@sct`
```{r}
test_error()
test_object("population_in_millions", undefined_msg = "Define the population first.", incorrect_msg = "Check code!")
test_function("hist",args=c("x"), incorrect_msg = "Check code again.")
success_msg("We got a histogram! Awesome!")
```

---
## Boxplots

Now we are going to make boxplots. Boxplots are useful when we want a summary of several variables or several strata of the same variables. Making too many histograms can become too cumbersome. 

```yaml
type: NormalExercise
key: 41b5550054
lang: r
xp: 100
skills: 1
```


`@instructions`
In one line of code:

- Stratify the state populations by region.
- Generate boxplots for the strate.

Note that you can achieve this using this `population~region` inside `boxplot` to generate the strata and specify the dataset with the `data` argument.

`@hint`
Make sure you specify the dataset. In the `boxplot` function specify the argument `data = murders`.

`@pre_exercise_code`
```{r}
library(dslabs)
```

`@sample_code`
```{r}
# Create a boxplot of state populations by region for the murders dataset

```

`@solution`
```{r}
# Create a boxplot of state populations by region and specify dataset 
boxplot(population~region, data = murders)

```

`@sct`
```{r}
test_error()
test_function("boxplot",args=c("formula","data"), incorrect_msg = "Check code. use ~ sign and include dataset.")
success_msg("Great job! Now you've learnt all three basic types of plots in R!")
```
---
## End of Assessment 8

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: bd1aeba6dc
```

This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/79c873ff79ea4680a67d98b169d2f2fc/21a56d47549740af97c3ef1ee0f868a6/2?activate_block_id=block-v1%3AHarvardX%2BPH125.1x%2B2T2017%2Btype%40vertical%2Bblock%40f9b9325c9f1f4f93b6bc8f557d8b0762'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K. WARNING: if you continue the assessments by clicking on the arrow to get the next exercise or by hitting Ctrl-K your assessments may NOT get scored.


`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```
