---
title       : Programming Basics
description : Programming Basics
---
## Conditionals

```yaml
type: MultipleChoiceExercise
key: 77691772c6
lang: r
xp: 100
skills: 1
```
What will this conditional expression return?
Run it from the console. 


```{r}
x <- c(1,2,-3,4)
if(all(x>0)){
    print("All Positives")
  } else{
     print("Not All Positives")
  }
```

`@instructions`
- All Positives
- Not All Positives
- N/A
- None of the above 

`@hint`
Are the numbers stored in `x` positive or not?

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Look at the values again."
msg2 = "Awesome! That is correct."
msg3 = "Check the question again."
msg4 = "Check the values in x once more."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Conditional continued

```yaml
type: MultipleChoiceExercise
key: 78dfb810dd
lang: r
xp: 50
skills: 1
```

Which of the following expressions is always `FALSE` when at least one entry of a logical vector x is `TRUE`?

`@instructions`
- all(x) 
- any(x) 
- any(!x) 
- all(!x)

`@hint`

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try again! Try an example on the R console."
msg2 = "Re-read the question! Try an example on the R console."
msg3 = "Check the question again. Try an example on the R console."
msg4 = "Good job! Let's move to the next question."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```
---
## ifelse

```yaml
type: NormalExercise
key: 04735d26b3
lang: r
xp: 100
skills: 1
```
The function `nchar` tells you how many characters long a character vector is. For example:

```{r}
library(dslabs)
data(murders)
char_len <- nchar(murders$state)
char_len[1:5]
```
`@instructions`
Use the `ifelse` function to write a line of code that assigns to the object new_names the state abbreviation when the state name is longer than 8 characters.

`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 

```

`@solution`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 
new_names <- ifelse(nchar(murders$state)>8, murders$abb, murders$state)

```

`@sct`
```{r}
test_error()
test_object("new_names", undefined_msg = "Make sure you define new_names!", incorrect_msg = "It has to be one line of code. Combine nchar and ifelse.")
success_msg("Woohoo! You're becoming a pro at this!")
```
---
## Defining functions

```yaml
type: NormalExercise
key: 3d324d0749
lang: r
xp: 100
skills: 1
```

You will encounter situations in which the function you need does not already exist. R permits you to write your own.  Let's practice one such situation, in which you first need to define the function to be used. The functions you define can have multiple arguments as well as default values.

`@instructions`
Create a function `sum_n` that for any given value, say `$n$`, computes the sum of the integers from 1 to n. Use the function to determine the sum of integers from 1 to 5,000.

`@hint`
- To make it inclusive, use {}. 
- Use code `sum_n <- function(n){sum(1:n)}` 

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create function called `sum_n`

# Use the function to determine the sum of integers from 1 to 5000

```

`@solution`
```{r}
# Create function called `sum_n`
sum_n <- function(n){sum(1:n)}

# Determine the sum of integers from 1 to 5000
sum_n(5000)

```

`@sct`
```{r}
test_error()
test_object("sum_n", undefined_msg = "Make sure you define sum_n!", incorrect_msg = "Use commands function(n) and sum.")
test_function("sum_n", incorrect_msg = "You need to check for the sum of 5000 integers!")
success_msg("This is awesome! Let's get to the next exercise.")
```
---
## Defining functions continued...

```yaml
type: NormalExercise
key: ea64484119
lang: r
xp: 100
skills: 1
```

`@instructions`

Create a function `altman_plot` that takes two arguments x and y and plots the difference(y-axis) against the sum(x-axis).


`@hint`
The command for plot, is `{plot(x+y, y-x)}`. 

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create `altman_plot` 

```

`@solution`
```{r}
# Create `altman_plot` 
altman_plot <- function(x, y){plot(x+y, y-x)}

```

`@sct`
```{r}
test_error()
test_object("altman_plot", undefined_msg = "Be sure to define altman_plot first.", incorrect_msg = "Make sure you are plotting (x,y) and calling the function that as well.")
success_msg("That's great! You can also play around with the variables and put in values for x and y.")
```
---
## Lexical scope

```yaml
type: NormalExercise
key: f13e10ed3c
lang: r
xp: 100
skills: 1
```

`@instructions`
```{r}

The lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which x is available when. After running the code below, what is the value of x?

```
`@hint`

`@pre_exercise_code`
```{r}
x <- 3
```

`@sample_code`
```{r}
# Run this code 
x <- 3
my_func <- function(y){
x <- 5
y+5
}

# Print value of x 
```

`@solution`
```{r}
# Print value of x
x
```

`@sct`
```{r}
test_error()
test_object("x", incorrect_msg = "How do you print the value of x?")
success_msg("Good job!") 
```
---
## For loops

```yaml
type: NormalExercise
key: efe45f5c56
lang: r
xp: 100
skills: 1
```

`@instructions`
We are soon going to write a `for` loop. In that `for` loop we are going to call a function. We define that function here. 
Write a function `compute_s_n` that for any given $n$ computes the sum `S_n = 1^2 + 2^2 + 3^2 +...+ n^2`. 
Report the value of the sum when `n=10`.

`@hint`

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Write the function for (n), with the above mentioned specifications and store it in `compute_s_n` 

# Report the value of the sum when n=10

```

`@solution`
```{r}
# Write the function for (n), with the above mentioned specifications and store it in `compute_s_n` 
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Report the value of the sum when n=10
compute_s_n(10)

```

`@sct`
```{r}
test_error()
test_object("compute_s_n", undefined_msg = "Store it in `compute_s_n`.", incorrect_msg = "Make sure you include both arguments in the function command.")
test_output_contains("compute_s_n(10)", incorrect_msg = "Place 10 in stead of n in the function.")
success_msg("That's awesome! You're getting so good at this!")
```
---
## For loops continued...

```yaml
type: NormalExercise
key: a9bf58a40c
lang: r
xp: 100
skills: 1
```
`@instructions`

Now define an empty numeric vector `s_n` of size 25 using `s_n <- vector("numeric", 25)` and store in the results of `S_1, S_2,...,S_25`.

`@hint`
Check the chapter on for loops for the codes!

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define a function and store it in `compute_s_n`


# Assign values to n and x 

```

`@solution`
```{r}
# Define a function and store it in `compute_s_n`
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Assign values to `n` and `s_n`
n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sct`
```{r}
test_error()
test_object("compute_s_n", undefined_msg = "Make sure to define compute_s_n first.", incorrect_msg = "Are you assigning the correct values to x? ")
test_object("n", incorrect_msg = "Check code.")
test_object("s_n", incorrect_msg = "Have you assigned values to n and x?")
success_msg("This is great! You now know the basics of for loops in R.")
```
---
## For loops and plotting

```yaml
type: NormalExercise
key: 4e649b5400
lang: r
xp: 100
skills: 1
```


`s_n` and `n` have already been defined in the previous question and are stored in memory for you!  

`@instructions`
Plot `s_n` versus `n`. Using points defined by `n=1,...,25`.

`@hint`
In the plot code, n is the x, and s_n is the y. 

`@pre_exercise_code`
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sample_code`
```{r}
#  Create the plot 

```

`@solution`
```{r}
#  Create the plot 
plot(n, s_n)

```

`@sct`
```{r}
test_error()
test_function("plot", incorrect_msg = "Did you plot the x and y axis correctly?")
success_msg("Awesome! Let's go to the last exercise in this chapter!")
```
---
## Identical

```yaml
type: NormalExercise
key: 33ed29e3c1
lang: r
xp: 100
skills: 1
```

`@instructions`
Confirm that `s_n = n(n+1)(2n+1)/6` using the `identical` command.

`@hint`

`@pre_exercise_code`
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sample_code`
```{r}
# Check that s_n is identical to the formula given in the instructions.

```

`@solution`
```{r}
# Check that s_n is identical to the formula given in the instructions.
identical(s_n, n*(n+1)*(2*n+1)/6)

```

`@sct`
```{r}
test_error()
test_output_contains("identical(s_n, n*(n+1)*(2*n+1)/6)", incorrect_msg = "Make sure you're checking your formula!")
success_msg("This is great! We are done with this module. Time to move on to bigger things!")
```



---

## End of Assessment 9

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: 226fb8de61
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/cfded5c208bc4e379606cb712cc54f25/5ba06674d0be41b99185b947e09e889b/?child=first'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```

