---
title                : R Basics 
description          : In this course we introduce you to the basics of computing and analyzing data in the user-friendly and helpful R interface. This first chapter starts with the very basics of functions, objects to get us acquainted with the world of R. 

---
## Using variables 1

```yaml
type: NormalExercise
key: 0cec712ce0
lang: r
xp: 100
skills: 1
```

What is the sum of the first $n$ positive integers? 

We can use the formula $ n(n+1)/2 $ to quickly compute this quantity.


`@instructions`
- Define 'n=100'
- Then use R to compute the sum of 1 through 100 using the formula $ n(n+1)/2$. What is the sum?

`@hint`
- Define `n` to be 100 in one line and simply type the formula using R code in the second line.
- Remember that in R you multiply using `*`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Here is how you compute the sum for the first 20 integers
20*(20+1)/2 

# However, we can define a variable to use the formula for other values of n
n <- 20
n*(n+1)/2

n <- 25
n*(n+1)/2

# Below, write your code to calculate the sum of the first 100 integers

```

`@solution`
```{r}
# Here is how you compute the sum for the first 20 integers
20*(20+1)/2 

# However, we can define a variable to use the formula for other values of n
n <- 20
n*(n+1)/2

n <- 25
n*(n+1)/2

# Below, write your code to calculate the sum of the first 100 integers 
n <- 100
n*(n+1)/2 
```

`@sct`
```{r}
# first instruction  
#test_object("n", incorrect_msg = "Make sure that you use `n` as your variable name and that you have assigned the correct value to `n`.")
#test_object("x", incorrect_msg = "Make sure that you use `n` as your variable name and that you have assigned the correct value to `n`.")
#test_output_contains ("12332424", incorrect_msg = "Take a look at your code for the second instruction.")
# General 
test_error() 
success_msg("Good job ! Let`s apply this to another question")
```
---
## Using variables 2

```yaml
type: NormalExercise
key: 213f198401
lang: r
xp: 100
skills: 1
```
What is the sum of the first $n$ positive integers? 

We can use the formula $ n(n+1)/2 $ to quickly compute this quantity.

`@instructions`
- Use the same formula as the last exercise but substitute `n`


`@hint`
Use the same R code as you used in the first question after changing the value of `n`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Below, write you code to calculate the sum of the first 1000 integers 

```

`@solution`
```{r}
# Below, write you code to calculate the sum of the first 1000 integers 
n <- 1000
n*(n+1)/2
```

`@sct`
```{r}
# test n 
test_object("n", incorrect_msg = "Something is wrong with `n`. Make sure you have assigned the correct value to `n`.")

# test correct output
test_output_contains ("(n*(n+1)/2)", incorrect_msg = "Take a look at your code for the second instruction.")

# General 
test_error() 
success_msg("Good job ! Let`s get to work on another question!")
```


---
## <<<New Exercise>>>

```yaml
type: MultipleChoiceExercise
key: 0e2a0ce37a
lang: r
xp: 100
skills: 1
```
Run the following code in the R console:

```{r, eval=FALSE}
n <- 1000
x <- seq(1,n)
sum(x)
```

Based on the result, what do you think the functions `seq` and `sum` do? You can use the help system.

`@instructions`
- sum creates a list of numbers and seq adds them up. 
- seq creates a list of numbers and sum adds them up. 
- seq computes the difference between two arguments and sum computes the sum of 1 through 1000. 
- sum always returns the same number

`@hint`
Go to R console and type `seq(1,5)`. See what you get. Then got the R console again and type `sum(seq(1,5))`. Change the `5` to other numbers.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try again! Read the choices carefully. Try again."
msg2 = "Well done. Proceed to the next exercise."
msg3 = "There`s no specificaton of 1 to 1000 for sum. Try again."
msg4 = "Read the help file for sum by typing `?sum` in the R console."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```
