---
title       : Basic Plots
description : We engage in the basics of plot-making in R, with scatterplots, histograms and box plots.
---
## Scatterplots

```yaml
type: NormalExercise
key: c535f6f10b
lang: r
xp: 100
skills: 1
```
We made a plot of total murders versus population and noted a strong relationship: not surprisingly states with larger populations had more murders. (Run the code in the console to get plot.)

```{r}
library(dslabs)
data(murders)
population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total
plot(population_in_millions, total_gun_murders)
```

Note that many states have populations below 5 million and are bunched up. We may gain further insights from making this plot in the log scale. 

`@instructions`
Transform the variables using the log10 transformation and then plot the total murders versus population.

`@hint`


`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Load the datasets and define some variables
library(dslabs)
data(murders)
population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total
plot(population_in_millions, total_gun_murders)

# Transform population using `log10` transformation

# Transform total gun murders using `log10` transformation

# Create a scatterplot with population and murders 

```

`@solution`
```{r}
# Transform population using `log10` transformation
log10_population<- log10(murders$population)

# Transform total gun murders using `log10` transformation
log10_total_gun_murders <- log10(murders$total)

# Create a scatterplot with population and murders 
plot(log10_population, log10_total_gun_murders)

```

`@sct`
```{r}
test_object("log10_population", undefined_msg = "Define log10_population first.", incorrect_msg = "Transform variable using log10.")
test_object("log10_total_gun_murders", undefined_msg = "Define log10_total_gun_murders first.", incorrect_msg = "Transform variable using log10.")
test_function("plot", incorrect_msg = "Scatter population on murders.")
test_error()
success_msg("Doesn't that plot look neat!")
```

---
## Histograms

```yaml
type: NormalExercise
key: 6d782c33a3
lang: r
xp: 100
skills: 1
```


`@instructions`
Create a histogram of the state populations.

`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Store the population in millions (same as previous question)
population_in_millions <- murders$population/10^6


# Create a histogram of the population 

```

`@solution`
```{r}
# Store the population in millions (same as previous question)
population_in_millions <- murders$population/10^6

# Create a histogram of the population 
hist(population_in_millions)

```

`@sct`
```{r}
test_error()
test_object("population_in_millions", undefined_msg = "Define the population first.", incorrect_msg = "Check code!")
test_function("hist", incorrect_msg = "Check code again.")
success_msg("We got a histogram! Awesome!")
```

---
## Boxplots

```yaml
type: NormalExercise
key: 41b5550054
lang: r
xp: 100
skills: 1
```


`@instructions`
Stratify the state populations by region, the generate boxplots for the strate.  You can achieve this using this `population~region` inside boxplot to generate the strata and specify the dataset with the `data` argument.

`@hint`
Make sure you specify the dataset.

`@pre_exercise_code`
```{r}
library(dslabs)
```

`@sample_code`
```{r}
# Create a boxplot of state populations by region and specify dataset 

```

`@solution`
```{r}
# Create a boxplot of state populations by region and specify dataset 
boxplot(population~region, data = murders)

```

`@sct`
```{r}
test_error()
test_function("boxplot", incorrect_msg = "Check code. use ~ sign and include dataset.")
success_msg("Great job! Now you've learnt all three basic types of plots in R!")

---

## End of Section 8

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/cfded5c208bc4e379606cb712cc54f25/5ba06674d0be41b99185b947e09e889b/?child=first'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```


```