---
title       : Programming Basics
description : Programming Basics
---
## Conditionals

```yaml
type: MultipleChoiceExercise
key: 77691772c6
lang: r
xp: 100
skills: 1
```
What will this conditional expression return?
Run it from the console. 

```{r}
x <- c(1,2,-3,4)
if(all(x>0)){
    print("All Positives")
  } else{
     print("Not All Positives")
  }
```

`@instructions`
- All Positives
- Not All Positives
- N/A
- None of the above 

`@hint`
Are the numbers stored in `x` all positive? If not the `print` after the `else` is evaluated.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Are all the entries in `x` positive?"
msg2 = "Awesome! That is correct."
msg3 = "In R you can vectorize logical relantions like this. There is no reason to have `NA`s"
msg4 = "If `all(x>0)` is `TRUE` or `FALSE` something is returned."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Conditional continued

```yaml
type: MultipleChoiceExercise
key: 78dfb810dd
lang: r
xp: 50
skills: 1
```

Which of the following expressions is always `FALSE` when at least one entry of a logical vector x is `TRUE`? You can try examples in the R console.

`@instructions`
- all(x) 
- any(x) 
- any(!x) 
- all(!x)

`@hint`
Define a logical vector `x` such as `c(TRUE, FALSE)` and `c(TRUE, TRUE)` and test them.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try `x <- c(TRUE, FALSE)` or `x <- c(TRUE, TRUE)`."
msg2 = "Try `x <- c(TRUE, TRUE)`"
msg3 = "Try `x <- c(TRUE, FALSE)"
msg4 = "Good job! Let's move to the next question."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```
---
## ifelse

```yaml
type: NormalExercise
key: 04735d26b3
lang: r
xp: 100
skills: 1
```
The function `nchar` tells you how many characters long a character vector is. For example:

```{r}
char_len <- nchar(murders$state)
head(char_len)
```

The function `ifelse` is useful because you convert a vectors of logicals into something else. For example, some datasets use the number -999 to denote NA. A bad practice! You can covert the -999 in a vector to NA using the following `ifelse` call:
```{r}
x <- c(2, 3, -999, 1, 4, 5, -999, 3, 2, 9)
ifelse(x == -999, NA, x)
```
If the entry is -999 it returns NA, otherwise it returns the entry.

`@instructions`
We will combine a number of functions for this exercise.
- Use the `ifelse` function to write one line of code that assigns to the object `new_names` the state abbreviation when the state name is longer than 8 characters.

`@hint`
Use a call to `nchar` as the first argument of the `ifelse` function to calculate. Check to see if it's larger than 8. Then you use `murders$abb` and `murders$state` in the second and third arguments to return either the abbreviation or state name depending on the number of characters.

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 

```

`@solution`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 
new_names <- ifelse(nchar(murders$state)>8, murders$abb, murders$state)

```

`@sct`
```{r}
test_error()
test_function("ifelse")
test_object("new_names", undefined_msg = "Make sure you define new_names!", incorrect_msg = "It has to be one line of code. Combine nchar and ifelse. The objects being used inside the function should be `murders$state` and `murders$abb`")
success_msg("Woohoo! You're becoming a pro at this!")

```
---
## Defining functions

```yaml
type: NormalExercise
key: 3d324d0749
lang: r
xp: 100
skills: 1
```

You will encounter situations in which the function you need does not already exist. R permits you to write your own.  Let's practice one such situation, in which you first need to define the function to be used. The functions you define can have multiple arguments as well as default values.

To define functions we use `function`. For example the following function adds 1 to the number it receives as an argument:

```{r}
my_func <- function(x){
    y <- x + 1
    y
}
```

The last value in the function, in this case that stored in `y`, gets returned.

If you run the code above R does not show anything. This means you defined the function. You can test it out like this:

```{r}
my_func(5)
```

`@instructions`
We will define a function `sum_n` for this exercise.
- Create a function `sum_n` that for any given value, say `n`, creates the vector `1:n`, and then computes the sum of the integers from 1 to n. 
- Use the function you just defined to determine the sum of integers from 1 to 5,000.

`@hint`
- To make it inclusive, use {}. 
- You can define the function like this:
```
sum_n <- function(n){
    x <- 1:n
    sum(x)
}
```

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create function called `sum_n`

# Use the function to determine the sum of integers from 1 to 5000

```

`@solution`
```{r}
# Create function called `sum_n`
sum_n <- function(n){
    x <- 1:n
    sum(x)
}

# Determine the sum of integers from 1 to 5000
sum_n(5000)

```

`@sct`
```{r}
test_error()
test_object("sum_n", undefined_msg = "Make sure you define sum_n!", incorrect_msg = "Use commands function(n) and sum.")
test_function("sum_n", incorrect_msg = "You need to check for the sum of 5000 integers!")
fun_def <- ex() %>% check_fun_def("sum_n")
fun_def %>% check_arguments()
fun_def %>% check_call(5000) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
success_msg("This is awesome! Let's get to the next exercise.")
```
---
## Defining functions continued...

We will make another function for this exercise. We will define a function `altman_plot` that takes two arguments `x` and `y` and plots the difference `y-x` in the y-axis against the sum `x+y` in the x-axis. 

You can define functions with as many variables as you want. For example, here we need at least two, `x` and `y`. The following function plot log transformed values:

```{r}
log_plot <- function(x, y){
    plot(log10(x), log10(y))
}
```

This function does not return anything. It just makes a plot.
```yaml
type: NormalExercise
key: ea64484119
lang: r
xp: 100
skills: 1
```

`@instructions`
We will make another function for this exercise.
- Create a function `altman_plot` that takes two arguments `x` and `y` and plots `y-x` versus `x+y`.

`@hint`
The command inside the function should be: `plot(x + y, y - x)`. Remember to define a function you use

```{r}
altman_plot <- function(var_1, var_2){
    # code that uses var_1 and var_2
}
```

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create `altman_plot` 

```

`@solution`
```{r}
# Create `altman_plot` 
altman_plot <- function(x, y){
    plot(x + y, y - x)
}
```

`@sct`
```{r}
test_error()
test_object("altman_plot", undefined_msg = "Be sure to define altman_plot first.", incorrect_msg = "Make sure you are plotting (x,y) and calling the function that as well.")
fun_def <- ex() %>% check_fun_def("altman_plot")
fun_def %>% check_arguments()
success_msg("That's great! You can also play around with the variables and put in values for x and y.")
```
---
## Lexical scope

Lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which `x` is available when.

```{r}
x <- 8
my_func <- function(y){
    x <- 9
    print(x)
    y + x
}
my_funct(x)
print(x)
```
Note that we `x` is 9, this is inside the function, but it is 8 after you run the function. The `x` changed inside the function but not outside. 

```yaml
type: NormalExercise
key: f13e10ed3c
lang: r
xp: 100
skills: 1
```

`@instructions`
The lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which `x` is available when. After running the code below, what is the value of `x`?
```{r}
x <- 3
my_func <- function(y){
    x <- 5
    y
    print(x)
}
my_funct(x)
```

`@hint`
Run the code and then `print(x)`. You will have the answer.

`@pre_exercise_code`
```{r}
x <- 3
```

`@sample_code`
```{r}
# Run this code 
x <- 3
    my_func <- function(y){
    x <- 5
    y+5
}

# Print value of x 
```

`@solution`
```{r}
# Show value of x
x
```

`@sct`
```{r}
test_error()
test_object("x", incorrect_msg = "How do you print the value of x?")
success_msg("Good job!") 
```
---
## For loops

In the next exercise we are going to write a for-loop. In that for-loop we are going to call a function. We define that function here. 

```yaml
type: NormalExercise
key: efe45f5c56
lang: r
xp: 100
skills: 1
```

`@instructions`
Write a function `compute_s_n` that for any given $n$ computes the sum $S_n = 1^2 + 2^2 + 3^2 + \dots + n^2$. 
Report the value of the sum when `n <- 10`.

`@hint`
What is it when you calculate by hand, does your function agree?

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Write the function for (n), with the above mentioned specifications and store it in `compute_s_n` 

# Report the value of the sum when n=10

```

`@solution`
```{r}
# Write the function for (n), with the above mentioned specifications and store it in `compute_s_n` 
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Report the value of the sum when n=10
compute_s_n(10)

```

`@sct`
```{r}
test_error()
test_object("compute_s_n", undefined_msg = "Store it in `compute_s_n`.", incorrect_msg = "Make sure you include both arguments in the function command.")
test_output_contains("compute_s_n(10)", incorrect_msg = "Place 10 instead of n in the function.")
fun_def <- ex() %>% check_fun_def("compute_s_n")
fun_def %>% check_arguments()
fun_def %>% check_call(10) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
success_msg("That's awesome! You're getting so good at this!")
```
---
## For loops continued...

```yaml
type: NormalExercise
key: a9bf58a40c
lang: r
xp: 100
skills: 1
```
`@instructions`
Storing values by for loops can be very useful.
Now define an empty numeric vector `s_n` of size 25 using `s_n <- vector("numeric", 25)` and store in the results of `S_1, S_2,...,S_25`.


`@hint`
Check the chapter on for loops for the codes!

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define a function and store it in `compute_s_n`

# Assign values to n and x 

```

`@solution`
```{r}
# Define a function and store it in `compute_s_n`
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Assign values to `n` and `s_n`
n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sct`
```{r}
test_error()
test_object("compute_s_n", undefined_msg = "Make sure to define compute_s_n first.", incorrect_msg = "Are you assigning the correct values to x? ")
test_object("n", incorrect_msg = "Check code.")
test_object("s_n", incorrect_msg = "Have you assigned values to n and x?")
test_function("vector")
fun_def <- ex() %>% check_fun_def("compute_s_n")
fun_def %>% check_arguments()
fun_def %>% check_call(10) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
for_loop <- ex() %>% check_for()
# check condition ("i in 1:5" in the solution)
# note that check_code might be too stringent, since someone could
# create a variable, e.g. a <- 1:5, and give that to the for loop
for_loop %>% check_cond() %>%
             check_code("in n", fixed = TRUE)
success_msg("This is great! You now know the basics of for loops in R.")
```
---
## For loops and plotting

```yaml
type: NormalExercise
key: 4e649b5400
lang: r
xp: 100
skills: 1
```


`s_n` and `n` have already been defined in the previous question and are stored in memory for you!  

`@instructions`
Using points defined by `n=1,...,25`.
Plot `s_n` versus `n`.

`@hint`
In the plot code, n is the x, and s_n is the y. 

`@pre_exercise_code`
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sample_code`
```{r}
#  Create the plot 

```

`@solution`
```{r}
#  Create the plot 
plot(n, s_n)

```

`@sct`
```{r}
test_error()
test_function("plot",args=c("x","y"), incorrect_msg = "Did you plot the x and y axis correctly?")
success_msg("Awesome! Let's go to the last exercise in this chapter!")
```
---
## Identical

```yaml
type: NormalExercise
key: 33ed29e3c1
lang: r
xp: 100
skills: 1
```

`@instructions`
Confirm that `s_n = n(n+1)(2n+1)/6` using the `identical` command.

`@hint`
`identical(s_n,n(n+1)(2n+1)/6)`

`@pre_exercise_code`
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

n <- 1:25
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

```

`@sample_code`
```{r}
# Check that s_n is identical to the formula given in the instructions.

```

`@solution`
```{r}
# Check that s_n is identical to the formula given in the instructions.
identical(s_n, n*(n+1)*(2*n+1)/6)

```

`@sct`
```{r}
test_error()
test_function("identical")
test_output_contains("identical(s_n, n*(n+1)*(2*n+1)/6)", incorrect_msg = "Make sure you're checking your formula!")
success_msg("This is great! We are done with this module. Time to move on to bigger things!")
```



---

## End of Assessment 9

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: 226fb8de61
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/6fd90a49a11d4ee3b94008b52f211653/f98acb1531854811b4c13907e7e55980/3?activate_block_id=block-v1%3AHarvardX%2BPH125.1x%2B2T2017%2Btype%40vertical%2Bblock%4080c3a1cbb4d64335bf1eb678472f122c'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```

