---
title       : Programming Basics
description : Programming Basics
---
## Conditionals

```yaml
type: MultipleChoiceExercise
key: 77691772c6
lang: r
xp: 100
skills: 1
```
What will this conditional expression return?
Run it from the console. 

```{r}
x <- c(1,2,-3,4)
if(all(x>0)){
    print("All Positives")
  } else{
     print("Not All Positives")
  }
```

`@instructions`
- All Positives
- Not All Positives
- N/A
- None of the above 

`@hint`
Are the numbers stored in `x` all positive? If not the `print` after the `else` is evaluated.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Are all the entries in `x` positive?"
msg2 = "Awesome! That is correct."
msg3 = "In R you can vectorize logical relations like this. There is no reason to have `NA`s"
msg4 = "If `all(x>0)` is `TRUE` or `FALSE` something is returned."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Conditional continued

```yaml
type: MultipleChoiceExercise
key: 78dfb810dd
lang: r
xp: 50
skills: 1
```

Which of the following expressions is always `FALSE` when at least one entry of a logical vector x is `TRUE`? You can try examples in the R console.

`@instructions`
- all(x) 
- any(x) 
- any(!x) 
- all(!x)

`@hint`
Define a logical vector `x` such as `c(TRUE, FALSE)` and `c(TRUE, TRUE)` and test them.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try `x <- c(TRUE, FALSE)` or `x <- c(TRUE, TRUE)`."
msg2 = "Try `x <- c(TRUE, TRUE)`"
msg3 = "Try `x <- c(TRUE, FALSE)"
msg4 = "Good job! Let's move to the next question."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```
---
## ifelse

```yaml
type: NormalExercise
key: 04735d26b3
lang: r
xp: 100
skills: 1
```
The function `nchar` tells you how many characters long a character vector is. For example:

```{r}
char_len <- nchar(murders$state)
head(char_len)
```

The function `ifelse` is useful because you convert a vector of logicals into something else. For example, some datasets use the number -999 to denote NA. A bad practice! You can convert the -999 in a vector to NA using the following `ifelse` call:
```{r}
x <- c(2, 3, -999, 1, 4, 5, -999, 3, 2, 9)
ifelse(x == -999, NA, x)
```
If the entry is -999 it returns NA, otherwise it returns the entry.

`@instructions`
We will combine a number of functions for this exercise.
- Use the `ifelse` function to write one line of code that assigns to the object `new_names` the state abbreviation when the state name is longer than 8 characters. So, for example, where the original vector has Massachusetts (13 characters), the new vector should have MA. But where the original vector has New York (8 characters), the new vector should have New York as well. 

`@hint`
Use a call to `nchar` as the first argument of the `ifelse` function to calculate. Check to see if it's larger than 8. Then you use `murders$abb` and `murders$state` in the second and third arguments to return either the abbreviation or state name depending on the number of characters.

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 


```

`@solution`
```{r}
# Assign the state abbreviation when the state name is longer than 8 characters 
new_names <- ifelse(nchar(murders$state)>8, murders$abb, murders$state)

```

`@sct`
```{r}
test_error()
test_function("ifelse")
test_object("new_names", undefined_msg = "Make sure you define new_names!", incorrect_msg = "It has to be one line of code. Combine nchar and ifelse. The objects being used inside the function should be `murders$state` and `murders$abb`. Also we are looking for `nchar(murders$state)` strictly larger than 8.")
success_msg("Woohoo! You're becoming a pro at this!")

```
---
## Defining functions

```yaml
type: NormalExercise
key: 3d324d0749
lang: r
xp: 100
skills: 1
```

You will encounter situations in which the function you need does not already exist. R permits you to write your own.  Let's practice one such situation, in which you first need to define the function to be used. The functions you define can have multiple arguments as well as default values.

To define functions we use `function`. For example the following function adds 1 to the number it receives as an argument:

```{r}
my_func <- function(x){
    y <- x + 1
    y
}
```

The last value in the function, in this case that stored in `y`, gets returned.

If you run the code above R does not show anything. This means you defined the function. You can test it out like this:

```{r}
my_func(5)
```

`@instructions`
We will define a function `sum_n` for this exercise.
- Create a function `sum_n` that for any given value, say `n`, creates the vector `1:n`, and then computes the sum of the integers from 1 to n. 
- Use the function you just defined to determine the sum of integers from 1 to 5,000.

`@hint`
- To make it inclusive, use {}. 
- You can define the function like this:

```{r}
sum_n <- function(n){
    x <- 1:n
    sum(x)
}
```

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create function called `sum_n`

# Use the function to determine the sum of integers from 1 to 5000

```

`@solution`
```{r}
# Create function called `sum_n`
sum_n <- function(n){
    x <- 1:n
    sum(x)
}

# Determine the sum of integers from 1 to 5000
sum_n(5000)

```

`@sct`
```{r}
test_error()
test_function("sum_n", incorrect_msg = "You need to check for the sum of 5000 integers!")
fun_def <- ex() %>% check_fun_def("sum_n")
fun_def %>% check_arguments()
fun_def %>% check_call(5000) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
success_msg("This is awesome! Let's get to the next exercise.")
```
---
## Defining functions continued...

We will make another function for this exercise. We will define a function `altman_plot` that takes two arguments `x` and `y` and plots the difference `y-x` in the y-axis against the sum `x+y` in the x-axis. 

You can define functions with as many variables as you want. For example, here we need at least two, `x` and `y`. The following function plots log transformed values:

```{r}
log_plot <- function(x, y){
    plot(log10(x), log10(y))
}
```

This function does not return anything. It just makes a plot.
```yaml
type: NormalExercise
key: ea64484119
lang: r
xp: 100
skills: 1
```

`@instructions`
We will make another function for this exercise.
- Create a function `altman_plot` that takes two arguments `x` and `y` and plots `y-x` (on the y-axis) against `x+y` (on the x-axis).

`@hint`
The command inside the function should be: `plot(x + y, y - x)`. Remember to define a function you use

```{r}
altman_plot <- function(var_1, var_2){
    # code that uses var_1 and var_2
}
```

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Create `altman_plot` 

```

`@solution`
```{r}
# Create `altman_plot` 
altman_plot <- function(x, y){
    plot(x + y, y - x)
}
```

`@sct`
```{r}
test_error()
test_function_definition("altman_plot",
                           body_test = {
                           	test_function("plot", incorrect_msg = "Make sure you call `plot` inside your function")
                            },
                         undefined_msg = "Be sure to define `altman_plot`.",
                         incorrect_number_arguments_msg = "Make sure the function takes two arguments.")
test_object("altman_plot", incorrect_msg = "Make sure you use `x` and `y` as your arguments and that you plot the difference `y-x` versus the sum `x+y`. Please write the sum as `x+y` not `y+x. Also, for this exercises, we want you to use the curly brackets `{ }` as in the example.") 
success_msg("That's great! You can also play around with the variables and put in values for x and y.")
```
---
## Lexical scope

Lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which `x` is available when.

```{r}
x <- 8
my_func <- function(y){
    x <- 9
    print(x)
    y + x
}
my_func(x)
print(x)
```
Note that when we define `x` as 9, this is inside the function, but it is 8 after you run the function. The `x` changed inside the function but not outside. 

```yaml
type: NormalExercise
key: f13e10ed3c
lang: r
xp: 100
skills: 1
```

`@instructions`
The lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which `x` is available when. After running the code below, what is the value of `x`?
```{r}
x <- 3
my_func <- function(y){
    x <- 5
    y
    print(x)
}
my_func(x)
```

`@hint`
Run the code and then `print(x)`. You will have the answer.

`@pre_exercise_code`
```{r}
x <- 3
```

`@sample_code`
```{r}
# Run this code 
x <- 3
    my_func <- function(y){
    x <- 5
    y+5
}

# Print value of x 
```

`@solution`
```{r}
# Show value of x
x
```

`@sct`
```{r}
test_error()
test_object("x", incorrect_msg = "How do you print the value of x?")
success_msg("Good job!") 
```
---
## For loops

In the next exercise we are going to write a for-loop. In that for-loop we are going to call a function. We define that function here. 

```yaml
type: NormalExercise
key: efe45f5c56
lang: r
xp: 100
skills: 1
```

`@instructions`
- Write a function `compute_s_n` that for any given $n$ computes the sum $S_n = 1^2 + 2^2 + 3^2 + \dots + n^2$. 
- Report the value of the sum when `n=10`.

`@hint`
The function will look something like this:

```{r}
compute_s_n <- function(n){
    x <- 1:n
    ## sum the vector after you square it
}
```
Then you evaluate it at 10: `compute_s_n(10)`

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Here is an example of function that adds numbers from 1 to n
example_func <- function(n){
    x <- 1:n
    sum(x)
}

# Here is the sum of the first 100 numbers
example_func(100)

# Write a function compute_s_n that with argument n and returns of 1 + 2^2 + ...+ n^2

# Report the value of the sum when n=10

```

`@solution`
```{r}
# Here is a function that adds numbers from 1 to n
example_func <- function(n){
    x <- 1:n
    sum(x)
}

# Here is the sum of the first 100 numbers
example_func(100)

# Write the function with argument n, with the above mentioned specifications and store it in `compute_s_n` 
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Report the value of the sum when n=10
compute_s_n(10)

```

`@sct`
```{r}
test_error()
test_function_result("compute_s_n", incorrect_msg ="Make sure your function is producing the correct output.")
test_output_contains("compute_s_n(10)", incorrect_msg = "Make sure your function is producing the correct output.")
test_student_typed("compute_s_n(10)", not_typed_msg = "Make sure you evaluate the function at 10 in the last line.")
fun_def <- ex() %>% check_fun_def("compute_s_n")
fun_def %>% check_arguments()
fun_def %>% check_call(10) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
success_msg("That's awesome! You're getting so good at this!")
```

---
## For loops continued...

Now we are going to compute the sum of the squares for several values of $n$. We will use a for-loop for this. Here is an example of a for-loop:

```{r}
results <- vector("numeric", 10)
n <- 10
for(i in 1:n){
    x <- 1:i
    results[i] <- sum(x)
}
```

Note that we start with a call to `vector` which constructs an empty vector that we will fill while the loop runs.

```yaml
type: NormalExercise
key: a9bf58a40c
lang: r
xp: 100
skills: 1
```
`@instructions`
- Define an empty numeric vector `s_n` of size 25 using `s_n <- vector("numeric", 25)`.
- Compute the the sum when n is equal to each integer from 1 to 25 using the function we defined in the previous exercise: `compute_s_n`
- Save the results in `s_n`


`@hint`
The line of code inside the function will be `s_n[i] <- compute_s_n(i)`. 

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define a function and store it in `compute_s_n`
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Create a vector for storing results
s_n <- vector("numeric", 25)

# write a for-loop to store the results in s_n

```

`@solution`
```{r}
# Define a function and store it in `compute_s_n`
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Create a vector for storing results
s_n <- vector("numeric", 25)

# Assign values to `n` and `s_n`
for(i in 1:25){
  s_n[i] <- compute_s_n(i)
}
```

`@sct`
```{r}
test_error()
test_object("compute_s_n", undefined_msg = "Make sure to define compute_s_n first.", incorrect_msg = "Are you assigning the correct values to x? ")
test_object("s_n", incorrect_msg = "You are getting the incorrect values. Check that you are calling the function correctly.")
test_function("vector")
fun_def <- ex() %>% check_fun_def("compute_s_n")
fun_def %>% check_arguments()
fun_def %>% check_call(10) %>% check_call(100) %>% check_call(1000)
fun_def %>% check_body()
for_loop <- ex() %>% check_for()
# check condition ("i in 1:5" in the solution)
# note that check_code might be too stringent, since someone could
# create a variable, e.g. a <- 1:5, and give that to the for loop
for_loop %>% check_cond() 
#%>% check_code("in n", fixed = TRUE)
success_msg("This is great! You now know the basics of for loops in R.")
```
---
## Checking our math

If we do the math, we can show that 

$$S_n = 1^2 + 2^2 + 3^2 + \dots + n^2 =  n(n+1)(2n+1)/6 $$

We have already computed the values of $S_n$ from 1 to 25 using a for loop. 

If the formula is correct then a plot of $S_n$ versus $n$ should look cubic.

Let's make this plot.
```yaml
type: NormalExercise
key: 4e649b5400
lang: r
xp: 100
skills: 1
```

`@instructions`
- Define `n <- 1:25`. Note that with this we can use `for(i in n)`
- Use a for loop to save the sums into a vector `s_n <- vector("numeric", 25)`
- Plot `s_n` (on the y-axis) against `n` (on the x-axis).

`@hint`
In the plot code, `n` is the x-axis, and `s_n` is the y-axis.

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define the function
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Define the vector of n
n <- 1:25

# Define the vector to store data
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

#  Create the plot 

```

`@solution`
```{r}
# Define the function
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Define the vector of n
n <- 1:25

# Define the vector to store data
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

#  Create the plot 
plot(n, s_n)
```

`@sct`
```{r}
test_error()
test_function("plot",args=c("x","y"), incorrect_msg = "Did you plot the x and y axis correctly?")
success_msg("Awesome! Let's go to the last exercise in this chapter!")
```
---
## Checking our math continued

Now let's actually check if we get the exact same answer.

```yaml
type: NormalExercise
key: 33ed29e3c1
lang: r
xp: 100
skills: 1
```

`@instructions`
- Confirm that `s_n` and $n(n+1)(2n+1)/6$ are the same using the `identical` command.

`@hint`
`s_n` is a vector. Because `n` is a vector `n*(n+1)*(2n+1)/6` is a vector. You can compare two vectors `x` and `y` with `identical(x, y)`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Define the function
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Define the vector of n
n <- 1:25

# Define the vector to store data
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

# Check that s_n is identical to the formula given in the instructions.

```

`@solution`
```{r}
# Define the function
compute_s_n <- function(n){
  x <- 1:n
  sum(x^2)
}

# Define the vector of n
n <- 1:25

# Define the vector to store data
s_n <- vector("numeric", 25)
for(i in n){
  s_n[i] <- compute_s_n(i)
}

# Check that s_n is identical to the formula given in the instructions.
identical(s_n, n*(n+1)*(2*n+1)/6)
```

`@sct`
```{r}
test_error()
test_function("identical")
test_output_contains("identical(s_n, n*(n+1)*(2*n+1)/6)", incorrect_msg = "Make sure you're checking your formula!")
success_msg("This is great! We are done with this module. Time to move on to bigger things!")
```



---

## End of Assessment 9

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: 226fb8de61
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/6fd90a49a11d4ee3b94008b52f211653/f98acb1531854811b4c13907e7e55980/3?activate_block_id=block-v1%3AHarvardX%2BPH125.1x%2B2T2017%2Btype%40vertical%2Bblock%4080c3a1cbb4d64335bf1eb678472f122c'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K. WARNING: if you continue the assessments by clicking on the arrow to get the next exercise or by hitting Ctrl-K your assessments may NOT get scored.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```

