---
title       : Vectors
description : The most basic unit to store data in R are vectors. We'll learn about creating and working with them in this chapter.
---
## Numeric Vectors

```yaml
type: NormalExercise
key: eb99436e3b
lang: r
xp: 100
skills: 1
```
A vector is series of values all of the same type. They are the most basic data type in R and can hold numeric data, character data, or logical data. In R, you can create a vector with the combine function `c()`. You place the vector elements separated by a comma between the parentheses. For example a numeric vector would look something like this: 

```{r}
cost <- c(50, 75, 90, 100, 150)
```

`@instructions`
Use the function `c()` to create a numeric vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto which are `35, 88, 42, 84, 81, and 30` degrees Fahrenheit. Call the object `temp`.

`@hint`
Make sure your numbers are within parenthesis and have commas separating them. 

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Here is an example creating a numeric vector named cost
cost <- c(50, 75, 90, 100, 150)

# Create a numeric vector to store the temperatures listed above into temp
# Make sure to follow the same order as above

```

`@solution`
```{r}
# Here is an example creating a numeric vector named cost
cost <- c(50, 75, 90, 100, 150)

# Create a numeric vector to store the temperatures listed above into temp
# Make sure to follow the same order as above
temp <- c(35, 88, 42, 84, 81, 30)
```

`@sct`
```{r}
test_error()
test_object("temp", incorrect_msg = "Check to match the numbers from the question!")
success_msg("Good job! Do you want to now try creating a character vector")
```