---
title       : Data Types
description : We're going to get an overview, of the different types of variables that there are in R, and how to work with them. 
---
## str

```yaml
type: MultipleChoiceExercise
key: 0e2a0ce37a
lang: r
xp: 100
skills: 1
```
We're going to be using the following dataset for this module. 
Run this code in the console. 

```{r}
library(dslabs)
data(murders)
```

Next, use the function `str` to examine the structure of the murders object. We can see that this object is a data frame with 51 rows and five columns. 

Which of the following best describes the variables represented in this data frame:


`@instructions`
- The 51 states 
- The murder rates for all 50 states and DC 
- The state name, the abbreviation of the state name, the state's region, and the state's population and the total number of murders for 2010.
- `str` shows no relevant information

`@hint`
Check the output that you see with the `str` function.

`@pre_exercise_code`
```{r}
library(dslabs)
str(murders) 
```

`@sct`
```{r}
msg1 = "Try again! Read the question again."
msg2 = "Try again! Read the question carefully."
msg3 = "Well done. Proceed to the next exercise"
msg4 = "Try again! Check what the question is asking for"
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```


