---
title       : Sorting
description : This chapter deals with getting more of an insight into the data we have, by helping us sort through it.
---
## sort

```yaml
type: NormalExercise
key: 8d2a5d166c
lang: r
xp: 100
skills: 1
```
When looking at a dataset, we may want to sort the data in an order that makes more sense for analysis. Let's learn to do this using the `murders` dataset as an example. 

`@instructions`
- Use the `$` operator to access the population size data and store it the object `pop`. 
- Then use the `sort` function to redefine `pop` so that it is sorted. 
- Finally use the `[` operator to report the smallest population size.

`@hint`
Make sure to use the `sort` function to redefine `pop` so that it is sorted.

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Access the `state` variable and store it in an object 
states <- murders$state 

# Sort the object alphabetically and redefine the object 
states <- sort(states) 

# Report the first alphabetical value  
states[1]

# Access `population` values from the dataset and store it in `pop`

# Sort the object and save it in the same object 

# Report the smallest population size 

```

`@solution`
```{r}
# Access the `state` variable and store it in an object 
states <- murders$state 

# Sort the object alphabetically and redefine the object 
states <- sort(states) 

# Report the first alphabetical value  
states[1]

# Access `population` values from the dataset and store it in `pop`
pop <- murders$population

# Sort the object and save it in the same object
pop <- sort(pop)

# Report the smallest population size 
pop[1]
```

`@sct`
```{r}
test_error()
test_object("pop", undefined_msg = "Make sure you define the object pop.", 
              incorrect_msg = "Check pop is properly defined and that you have redefined pop using the sort function.")
test_function("sort", index = 2,eval = TRUE,eq_condition = "equivalent", not_called_msg = "Make sure to call the sort function.",
              incorrect_msg = "Make sure to call it on the object `pop`.")
test_output_contains("pop[1]", incorrect_msg = "Use the square brackets [] to report the smallest population value.")
success_msg("Good job! Now you know how to sort data in an ascending order.")
```

---
## order

```yaml
type: NormalExercise
key: 9bf4a9a8ca
lang: r
xp: 100
skills: 1
```


`@instructions`
Now instead of the smallest population size, let's find out the index number of the state with the smallest population size, using the command `order`. 

`@hint`
Replace sort with order

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Access `population` from the dataset and store it in `pop`

# Use the command `order`, to order `pop` and store in object `o`

# Find the index number of the entry with the smallest population size

```

`@solution`
```{r}
# Access `population` from the dataset and store it in `pop`
pop <- murders$population

# Use the command `order`, to order `pop` and store in object `o`
o <- order(pop)

# Find the index number of the entry with the smallest population size
o[1]

```

`@sct`
```{r}
test_error()
test_object("pop", undefined_msg = "Define pop first.", incorrect_msg = "Make sure you save population in pop.")
test_object("o", undefined_msg = "Make sure you define o first!", incorrect_msg = "Store pop using the order code in object o.")
test_output_contains("o[1]", incorrect_msg = "Use the command provided in the instructions.") 
test_function("order", index = 1,eval = TRUE,eq_condition = "equivalent", not_called_msg = "Make sure to call the order function.",
              incorrect_msg = "Make sure to call it on the object `pop`.")
success_msg("Great job!")
```

---
## New Codes 

We can actually perform the same operation as in the previous exercise using the function `which.min`. It basically tells us which is the minimum value. 

```yaml
type: NormalExercise
key: 69b30d5992
lang: r
xp: 100
skills: 1
```


`@instructions`
Write one line of code that gives the index of the lowest population entry, using the `which.min` command.

`@hint`
Use `which.min` directly. 

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Find the smallest value for variable total 
which.min(murders$total)

# Find the smallest value for `population`

```

`@solution`
```{r}
# Find the smallest value for variable total 
which.min(murders$total)

# Find the smallest value for `population`
which.min(murders$population)

```

`@sct`
```{r}
test_error()
test_output_contains("which.min(murders$population)", incorrect_msg = "Make sure you to use $ to get the population data and simultaneously use which.min.")
test_function("which.min",index=2)
test_student_typed("which.min(murders$population)",
                   fixed = TRUE,times = 1,not_typed_msg ="Use the $ simultaneously with the which.min function.")
success_msg("Great ! You've learnt another code in R!")
```

---
## Using the output of order 

Now we know how small the smallest state is and we know which row represents it. However, which state is it? 

```yaml
type: NormalExercise
key: 3daf04cc14
lang: r
xp: 100
skills: 1
```


`@instructions`
Define a variable `states` to hold the state names from the murders data frame. Find the index of the smalles state using `which.min(murders$population)]`. Then combine these to find the state name for the smallest state.

`@hint`
Index the list of variables

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Define the variable `pop` to hold the population sizes

# Define the variable `i` to be the index of the smallest state

# Define variable `state` to hold the states

# Use the index you define to find the state with the smallest population
```

`@solution`
```{r}
# Define the variable `pop` to hold the population sizes
pop <- murders$population

# Define the variable `i` to be the index of the smallest state
i <- which.min(pop)

# Define variable `states` to hold the states
states <- murders$state

# Use the index you define to find the state with the smallest population
states[i]
```

`@sct`
```{r}
test_error()
test_object("states", undefined_msg = "Define states first!", incorrect_msg = "Assign state values from dataset to states.")
test_output_contains("states[which.min(murders$population)]", incorrect_msg = "Copy code from instructions.")
success_msg("Awesome! Now we have the names of the states as well!")
```

---
## Ranks

You can create a data frame using the `data.frame` function. Here is a quick example:

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name = city, temperature = temp)
```

```yaml
type: NormalExercise
key: 655d8a25c3
lang: r
xp: 100
skills: 1
```


`@instructions`
Use the `rank(murders$population)` function to determine the population size rank (from smallest to biggest) of each state. 
Save these ranks in an object called `ranks`.
Then create a data frame with the state name and its rank. Call the data frame `my_df`. 


`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Store temperatures in an object 
temp <- c(35, 88, 42, 84, 81, 30)

# Store city names in an object 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Create data frame with city names and temperature 
city_temps <- data.frame(name = city, temperature = temp)

# Define a variable states to be the state names 


# Define a variable `ranks` to determine the population size ranks 


# Create a dataframe `my_df` with the state name and its rank

```

`@solution`
```{r}
# Define a variable states to be the state names from the murders data frame
states <- murders$state

# Define a variable `ranks` to determine the population size ranks 
ranks <- rank(murders$population)

# Create a dataframe `my_df` with the state name and its rank
my_df <- data.frame(states = states, ranks = ranks)
```

`@sct`
```{r}
test_error()
test_object("states", incorrect_msg = "Store the state names in states!")
test_object("ranks", undefined_msg = "Define ranks first!", incorrect_msg = "Define the rank of the population values!")
test_object("my_df", undefined_msg = "Define the dataframe first.", incorrect_msg = "Use the command similar to the example.")
success_msg("That's awesome! You got this!")
```

---
## Data Frames, Ranks and Orders 

```yaml
type: NormalExercise
key: 62209cebec
lang: r
xp: 100
skills: 1
```


`@instructions`
This exercise is somewhat more challenging. We are going to 
repeat the previous exercise but this time order `my_df` so that the states are ordered from least populous to most.
Create an object ind that stores the indexes needed to order the population values, using the `order` command.  For example we could define `o <- order(murders$population)`
Then use the bracket operator `[` to re-order each column in the data frame. For example, `abb[o]` orders the abbreviations based by population size. 

`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Define a variable states to be the state names from the murders data frame


# Define a variable `ranks` to determine the population size ranks 


# Define a variable `ind` to store the indexes needed to order the population values


# Create a dataframe `my_df` with the state name and its rank and ordered from least populous to most 

```

`@solution`
```{r}
# Define a variable states to be the state names from the murders data frame
states <- murders$state

# Define a variable `ranks` to determine the population size ranks 
ranks <- rank(murders$population)

# Define a variable `ind` to store the indexes needed to order the population values
ind <- order(states)

# Create a dataframe `my_df` with the state name and its rank and ordered from least populous to most 
my_df <- data.frame(states = states[ind], ranks = ranks[ind])

```

`@sct`
```{r}
test_error()
test_object("states", undefined_msg = "Define states first.", incorrect_msg = "")
test_object("ranks",undefined_msg = "Define ranks first.", incorrect_msg = "")
test_object("ind",undefined_msg = "Define ind first.", incorrect_msg = "Use the order command as well.")
test_object("my_df", undefined_msg = "Define my_df first", incorrect_msg = "Check code")
success_msg("Great job! See how well you're building on your knowledge of R!")
```

---
## NA

The `na_example` dataset represents a series of counts. You can quickly examine the object using

```{r}
library(na_example)
data(na_example)
str(na_example)
```

However, when we compute the average we obtain an `NA`. You can see this by typing 

```{r}
mean(na_example)
```

```yaml
type: NormalExercise
key: babfb93aab
lang: r
xp: 100
skills: 1
```


`@instructions`
The `is.na` returns a logical vector that tells us which entries are `NA`.
Assign this logical vector to an object called `ind`, using `is.na(na_example)` and determine how many NA does `na_example` have, using the `sum` command.

`@hint`
Sum the indices

`@pre_exercise_code`
```{r}
library(dslabs)
data(na_example)  
```

`@sample_code`
```{r}
# Using new dataset 
library(dslabs)
data(na_example)

# Checking the structure 
str(na_example)

# Find out the mean of the entire dataset 
mean(na_example)

# Use `is.na` to create an logical index `ind` that tells which entries are NA

# Determine how many NA does `ind` have

```

`@solution`
```{r}
# Assign `is.na` to logical vector `ind` 
ind <- is.na(na_example)

# Determine how many NA does `ind` have? 
sum(ind)

```

`@sct`
```{r}
test_error()
test_object("ind", undefined_msg = "Make sure to define ind first.", incorrect_msg = "Check instructions for the code.")
test_output_contains("sum(ind)", incorrect_msg = "Use the sum command to get the number of NAs.")
success_msg("Great job! Now let's move to one last thing in this chapter.")
```

---
## Removing NAs 

A common operation we perform is removing the entries that are NA and do operations on the rest. 


```yaml
type: NormalExercise
key: 0462684851
lang: r
xp: 100
skills: 1
```


`@instructions`
Compute the average again but only for the entries that are not NA making use of the `!` operator before `ind`. 

`@hint`
Remember the `!` operator. Use code `mean(na_example[!ind])`. 

`@pre_exercise_code`
```{r}
library(dslabs)
data(na_example)
ind <- is.na(na_example)
```

`@sample_code`
```{r}
# Compute the average, for entries that are not NA 

```

`@solution`
```{r}
# Compute the average, for entries that are not NA 
mean(na_example[!ind])

```

`@sct`
```{r}
test_error()
test_output_contains("mean(na_example[!ind])", incorrect_msg = "Check your code from the instructions again.")
success_msg("Awesome ! Now you're all set with vectors! Get practicing on your own!")
```
