---
title       : Data Types
description : We're going to get an overview, of the different types of variables that there are in R, and how to work with them. 
---
## str

```yaml
type: MultipleChoiceExercise
key: 0e2a0ce37a
lang: r
xp: 100
skills: 1
```
We're going to be using the following dataset for this module. 
Run this code in the console. 

```{r}
library(dslabs)
data(murders)
```

Next, use the function `str` to examine the structure of the murders object. We can see that this object is a data frame with 51 rows and five columns. 

Which of the following best describes the variables represented in this data frame:


`@instructions`
- The 51 states 
- The murder rates for all 50 states and DC 
- The state name, the abbreviation of the state name, the state's region, and the state's population and the total number of murders for 2010.
- `str` shows no relevant information

`@hint`
Check the output that you see with the `str` function.

`@pre_exercise_code`
```{r}
library(dslabs)
str(murders) 
```

`@sct`
```{r}
msg1 = "Try again! Read the question again."
msg2 = "Try again! Read the question carefully."
msg3 = "Well done. Proceed to the next exercise"
msg4 = "Try again! Check what the question is asking for"
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Variable names

```yaml
type: NormalExercise
key: 7455b7cb2b
lang: r
xp: 100
skills: 1
```
In the previous question, we saw the different variables that are a part of this dataset from the output of the `src()` function. The  function `names()` is specifically designed to extract the column names from a data frame. 

`@instructions`
What are the column names used by the murders data frame for the five variables?

`@hint`
Make sure you put the name of the dataset - `murders` - in parenthesis after names. 

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Load package and data

library(dslabs)
data(murders)

# Checking the column names 

```

`@solution`
```{r}
# Load package and data

library(dslabs)
data(murders)

# Checking the column names 
names(murders)
```

`@sct`
```{r}
test_error()
test_function("names", incorrect_msg = "Make sure you've put in the name of the dataset.") 
success_msg("Good job!")
```
