---
title       : Vector Arithmetics
description : We will learn about the more powerful vector arithmetic functions in R, that help us solve higher-level problems. 
---
## Vectorized operations

```yaml
type: NormalExercise
key: 496010199b
lang: r
xp: 100
skills: 1
```
Previously we created this data frame:

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name = city, temperature = temp)
```

`@instructions`
We are interested in specifying Celsius instead of Fahrenheit. Remember that to convert from Fahrenheit to Celsius we use $C = 5/9 \times (F-32)$.

- Use vector arithmetics to convert `temp` to Celsius
- Create a data frame called `city_temps` with the city names and temperatures in Celsius.

`@hint`
To subtract a constant `a` to each entry of a vector `x` and then multiply each by `b` to each use `(x-a)*b`

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Assign city names to `city` 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Store temperature values in `temp`
temp <- c(35, 88, 42, 84, 81, 30)

# Convert temperature into Celsius and overwrite the original values of 'temp' with these Celsius values

# Create a data frame `city_temps` 

```

`@solution`
```{r}
# Assign city names to `city` 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Store temperature values in `temp`
temp <- c(35, 88, 42, 84, 81, 30)

# Convert temperature into Celsius and overwrite the original values of 'temp' with these Celsius values
temp <- (temp - 32)*5/9

# Create data frame `city_temps` 
city_temps <- data.frame(name = city, temperature = temp)

```

`@sct`
```{r}
test_error()
test_object("temp", incorrect_msg = "Change from Fahrenheit to Celsius using `temp <- (temp - 32)*5/9`.")
test_object("city", incorrect_msg = "Store names of cities.")
test_object("city_temps", incorrect_msg = "Create a data frame following the same example.")
success_msg("Great job!")
```

---
## Vectorized operations continued...

We can use some of what we have learned to perform calculations that 
would otherwise be quite complicated. Let's see an example.


```yaml
type: NormalExercise
key: 657f51ea05
lang: r
xp: 100
skills: 1
```


`@instructions`

What is the sum of the following equation: `1 + 1/2^2 + 1/3^2 + ... + 1/100^2`? Thanks to Euler we know it should be close to $\pi^2/6$. 

- Define an object `x` that contains the numbers 1 through 100
- Compute the sum `1 + 1/2^2 + 1/3^2 + ... + 1/100^2`.

`@hint`
Remember that you can use vectorization to get the series using `1/x^2`. You can just sum these vectors.

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define an object `x` with the numbers 1 through 100

# Compute the sum 

```

`@solution`
```{r}
# Define an object `x` with the numbers 1 through 100
x <- seq(1, 100)

# Sum the equation 
sum(1/x^2) 
```

`@sct`
```{r}
test_error()
test_object("x", undefined_msg = "Define x first!", incorrect_msg = "Use the sequence command to store 1,100 in x.")
test_function("sum", incorrect_msg = "Use the code given in the instructions.")
test_output_contains("sum(1/x^2)")
success_msg("Good job!")
```

---
## Vectorized operation continued...

```yaml
type: NormalExercise
key: e5215311fc
lang: r
xp: 100
skills: 1
```


`@instructions`
- Compute the per 100,000 murder rate for each state and store it in the object `murder_rate`. 
- Then compute the average of the state murder rates for the US using the function `mean`. What is this average?

`@hint`
You can use vectorization to compute the rate. If `tot` has the total murders and `pop` has the population for each state, then `tot/pop*100000` has the rate each state. Then you can simply compute the average using `mean`.

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Load the data
library(dslabs)
data(murders)

# Store the per 100,000 murder rate for each state in murder_rate

# Calculate the average murder rate in the US 

```

`@solution`
```{r}
# Load the data
library(dslabs)
data(murders)

# Store the per 100,000 murder rate for each state in murder_rate
murder_rate <- murders$total/murders$population*100000

# Calculate the average murder rate in the US
mean(murder_rate)

```

`@sct`
```{r}
test_error()
test_object("murder_rate", incorrect_msg = "")
test_function("mean", incorrect_msg = "")
test_output_contains("mean(murder_rate)",incorrect_msg = "Make sure to multiply the population by 100,000")
success_msg("Awesome!")
```

---

## End of Assessment 5

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: 9831da4a5f
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/2a638f36b59f4897b1bd6867ad58f63e/63b7518d4edc47518dbe29c1e5db8822/2?activate_block_id=block-v1%3AHarvardX%2BPH125.1x%2B2T2017%2Btype%40vertical%2Bblock%405d7cd083d32843e39f547864236f966b'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K. WARNING: if you continue the assessments by clicking on the arrow to get the next exercise or by hitting Ctrl-K your assessments may NOT get scored.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```

