---
title       : Vector Arithmetics
description : We will learn about the more powerful vector arithmetic functions in R, that help us solve higher-level problems. 
---
## Vectorized operations

```yaml
type: NormalExercise
key: 496010199b
lang: r
xp: 100
skills: 1
```
Previously we created this data frame:

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name = city, temperature = temp)
```

`@instructions`
- Remake the data frame using the code above but add a line that converts the temperature from farenheit to celsius. 

`@hint`
The conversion rate is: `degree Farenheit* 9/5 + 32 = degree celsius` 

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Store temperature values in `temp`
temp <- c(35, 88, 42, 84, 81, 30)

# Convert temperature into celsius and restore in `temp`


# Assign city names to `city` 


# Create a data frame `city_temps` 

```

`@solution`
```{r}
# Store temperature values in `temp`
temp <- c(35, 88, 42, 84, 81, 30)

# Convert temperature into celsius and restore in `temp`
temp <- (temp - 32)*5/9

# Assign city names to `city` 
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Create data frame `city_temps` 
city_temps <- data.frame(name = city, temperature = temp)

```

`@sct`
```{r}
test_object("temp", undefined_msg = "Define temp first.", incorrect_msg = "Store values of temperature!")
test_object("temp", incorrect_msg = "Change from Farenheit to Celsius using `temp <- (temp - 32)*5/9`.")
test_object("city", incorrect_msg = "Store names of cities.")
test_object("city_temps", incorrect_msg = "Create a dataframe following the same example.")
test_error()
success_msg("Great job!")
```

---
## Vectorized operations continued

We can use some of what we have learned to perform calculations that 
would otherwise be quite complicated. Let's see an example.


```yaml
type: NormalExercise
key: 657f51ea05
lang: r
xp: 100
skills: 1
```


`@instructions`
What is the sum of the following equation: `1 + 1/2^2 + 1/3^2 + ... + 1/100^2`? Use command `sum(1/x^2)`.
Thanks to Euler we know it should be close to $\pi^2/6$. 

`@hint`
This is a series.

`@pre_exercise_code`
```{r}
# no pec 
```

`@sample_code`
```{r}
# Define an object `x` with the numbers 1 through 100

# Sum the equation 

```

`@solution`
```{r}
# Define an object `x` with the numbers 1 through 100
x <- seq(1, 100)

# Sum the equation 
sum(1/x^2) 
```

`@sct`
```{r}
test_error()
test_object("x", undefined_msg = "Define x first!", incorrect_msg = "Use the sequence command to store 1,100 in x.")
test_function("sum", incorrect_msg = "Use the code given in the instructions.")
success_msg("Good job!")
```

---
## Vectorized operation continued

```yaml
type: NormalExercise
key: e5215311fc
lang: r
xp: 100
skills: 1
```


`@instructions`
Compute the per 100,000 murder rate for each state and store it in the object `murder_rate`. 
Then compute the average murder rate for the US using the function `mean`. What is the average?

`@hint`

`@pre_exercise_code`
```{r}
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Load the data
library(dslabs)
data(murders)

# Store the per 100,000 murder rate for each state in `murder_rate`

# Calculate the average murder rate in the US 

```

`@solution`
```{r}
# Load the data
library(dslabs)
data(murders)

# Store the per 100,000 murder rate for each state in `murder_rate`
murder_rate <- murders$total/murders$population*100000

# Calculate the average murder rate in the US
mean(murder_rate)

```

`@sct`
```{r}
test_error()
test_object("murder_rate", incorrect_msg = "")
test_function("mean", incorrect_msg = "")
success_msg("Awesome!")
```

---

## End of Section 4

```yaml
type: PlainMultipleChoiceExercise
lang: r
xp: 0
skills: 1
key: c389117b79
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/cfded5c208bc4e379606cb712cc54f25/5ba06674d0be41b99185b947e09e889b/?child=first'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```

