---
title       : Basic Data Manipulation 
description : Just as the name suggests, we're going to practice the basics of data manipulation here.
---
## dplyr

```yaml
type: NormalExercise
key: 766d9c4eab
lang: r
xp: 100
skills: 1
```
Load the `dplyr` package and the murders dataset.

```{r}
library(dplyr)
library(dslabs)
data(murders)
```

`@instructions`
Use the function `mutate` to add a murders rate column with the per 100,000 murder rate.

`@hint`
Use code `murders <- mutate(murders, rate =  total / population * 100000)`.  

`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Use mutate to add a column named rate and store in murders

```

`@solution`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Use mutate to add a rate column and store in `murders`
murders <- mutate(murders, rate =  total / population * 100000)
```

`@sct`
```{r}
test_object("murders", undefined_msg = "Define murder first.", incorrect_msg = "Check the code again.")
test_error()
success_msg("Awesome! Let's learn another command with mutate.")
```

---
## Mutate

```yaml
type: NormalExercise
key: 79bec8ece0
lang: r
xp: 100
skills: 1
```


`@instructions`
Use the function `mutate` again to add a column `rank` containing the rank, from highest to lowest murder rate. Note that if `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest.

`@hint`
Use the function rank and remember by default it ranks form lowest to highest.

`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
rate =  murders$total)/ (murders$population) * 100000
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Defining rate
rate <-  murders$total/ murders$population * 100000

# Use mutate to add a rate column, rank from highest to lowest and store in `murders`

```

`@solution`
```{r}

# Use mutate to add a rate column, rank from highest to lowest and store in `murders`
murders <- mutate(murders, rank = rank(-rate))
```

`@sct`
```{r}
test_error()
test_object("murders", incorrect_msg = "Remember, rank from highest to lowest.")
success_msg("Good job!")
```
---
## filter

```yaml
type: NormalExercise
key: 9a56fa2057
lang: r
xp: 100
skills: 1
```


`@instructions`

The `dplyr` function `filter` is used to select specific rows of the data frame to keep. Show the top 5 states with the highest murder rates. Do not change the murders dataset, just show the result.

`@hint`


`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)

```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Add the necessary columns
murders <- mutate(murders, rate = total/population * 100000, rank = rank(-rate))

# Filter to show the top 5 states with the highest murder rates

```

`@solution`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Add the necessary columns
murders <- mutate(murders, rate = total/population * 100000, rank = rank(-rate))

# Filter to show the top 5 states with the highest murder rates
filter(murders, rank <= 5)
```

`@sct`
```{r}
test_error()
test_output_contains("filter(murders, rank <= 5)", incorrect_msg = "Make sure you've written murders in your code.")
success_msg("Great job!")
```
---
## filter with !=

```yaml
type: NormalExercise
key: 846dd29c6c
lang: r
xp: 100
skills: 1
```


`@instructions`
Create a new data frame called `no_south` that removes states from the south. How many states are in this category? 

`@hint`
Use filter and the != operator. You can use nrow to quickly get the number of rows of a data frame.

`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Use filter Creat a new dataframe `no_south`

# Use nrow() to calculate the number of rows
```

`@solution`
```{r}
# Creating a new dataframe `no_south`
no_south <- filter(murders, region != "South" )

# Number of states (rows) in this category 
nrow(no_south)

```

`@sct`
```{r}
test_error()
test_object("no_south", undefined_msg = "Define no_south first!", incorrect_msg = "Make sure you`re excluding the ones not in the South.")
test_output_contains("nrow(no_south)", incorrect_msg = "You need the number of rows for the number of states.")
success_msg("That's great! Let's move to the next exercise.")
```

---
## filter with %in%

```yaml
type: NormalExercise
key: 4e5d7ac60f
lang: r
xp: 100
skills: 1
```


`@instructions`
Create a new dataframe called murders_nw with only the states from the northeast and the west. 
How many states are in this category? 


`@hint`
Use filter and the `%in%` operator. You can use nrow to quickly get the number of rows of a data frame.

`@pre_exercise_code`
```{r}
library(dplyr)
libary(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Create a new dataframe called `murders_nw` with only the states from the northeast and the west


# Number of states (rows) in this category 

```

`@solution`
```{r}
# Create a new dataframe called `murders_nw` with only the states from the northeast and the west
murders_nw <- filter(murders, region %in% c("Northeast", "West") )

# Number of states (rows) in this category
nrow(murders_nw)

```

`@sct`
```{r}
test_error()
test_object("murders_nw", undefined_msg = "Define murders_nw first.", incorrect_msg = "Make sure you're using the %in% command.")
test_output_contains("nrow(murders_nw)", incorrect_msg = "You need the number of rows again, same as previous question.")
success_msg("This is great!")
```
---
## filtering by two conditions

```yaml
type: NormalExercise
key: 6e8611adb6
lang: r
xp: 100
skills: 1
```


`@instructions`
Suppose you want to live in the Northeast or West and want the murder rate to be less than 1. 
Create a table, call it `my_states`, that satisfies both these. 
How many states satisfy this?

`@hint`
Use the `my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)` code. 

`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# add the rate column
murders <- mutate(murders, rate =  total / population * 100000)

# Create a table, call it my_states, that satisfies both the conditions 

# Number of states (rows) in this category

```

`@solution`
```{r}
# Create a table, call it `my_states`, that satisfies both the conditions 
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)

# Number of states (rows) in this category
nrow(my_states)

```

`@sct`
```{r}
test_error()
test_object("my_states", undefined_msg = "Define my_states first!", incorrect_msg = "Use the filter, %in% and < commands.")
test_output_contains("nrow(my_states)", incorrect_msg = "You need the number of rows again, same as previous question.")
success_msg("Now you know how to combine functions and use them to get specific answers. Let's try it one more time!")
```
---
## Using the pipe %>%

```yaml
type: NormalExercise
key: 72305c532e
lang: r
xp: 100
skills: 1
```


`@instructions`
Repeat exercise 5, but now instead of creating a new obejct, show the result and only include the state, rate, and rank columns. Use a pipe (`%>%`) to do this in just one line.

`@hint`
Use code: `filter(murders, region %in% c("Northeast", "West") & rate < 1) %>% 
    select(state, rate, rank)` 
    
`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
rate =  (murders$total)/ (murders$population) * 100000
```

`@sample_code`
```{r}
## Define the rate column
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))

# show the result and only include the state, rate, and rank columns, all in one line

```
`@solution`
```{r}
## Define the rate and rank column
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))

# show the result and only include the state, rate, and rank columns, all in one line
filter(murders, region %in% c("Northeast", "West") & rate < 1) %>%  
   select(state, rate, rank)
   
```
`@sct`
```{r}
test_error()
test_function("filter", incorrect_msg = "Everything needs to be in one line! Use a pipe to help!")
success_msg("You're getting a pretty good hang of stuff now!")
```
---
## mutate, filter and select

```yaml
type: NormalExercise
key: 60fa0dbe3a
lang: r
xp: 100
skills: 1
```


`@instructions`
Now reload the original murders data frame using `data(murders)`. Then use just one line to create a new data frame, called, `my_states` that has murder rate and rank column, consideres only states in the Northeast or West, has a murder rate lower than 1 and contains only the state, rate, and rank columns. 

`@hint`
Use `mutate`, `filter` and `select` combining them with `pipes`.

`@pre_exercise_code`
```{r}
library(dplyr)
library(dslabs)
data(murders)
```

`@sample_code`
```{r}
# Loading the libraries
library(dplyr)
data(murders)

# Create new data frame called `my_states` (with specifications in the instructions)

```

`@solution`
```{r}
# Create new data frame called `my_states` (with specifications in the instructions)
my_states <- murders %>% 
    mutate(rate =  total / population * 100000, rank = rank(-rate)) %>%
    filter(region %in% c("Northeast", "West") & rate < 1) %>%
    select(state, rate, rank)
    
```

`@sct`
```{r}
test_error()
test_object("my_states", undefined_msg = "Define my_states!", incorrect_msg = "Make sure you put pipes to put everything in one line.")
success_msg("This is absolutely awesome! You now know how to use basic data manipulation techniques in R.")
```
