---
title                : R Basics 
description          : In this course we introduce you to the basics of computing and analyzing data in the user-friendly and helpful R interface. This first chapter starts with the very basics of functions, objects to get us acquainted with the world of R. 

---
## Using variables 1

```yaml
type: NormalExercise
key: 0cec712ce0
lang: r
xp: 100
skills: 1
```

What is the sum of the first $n$ positive integers? 

We can use the formula $ n(n+1)/2 $ to quickly compute this quantity.


`@instructions`
- Define 'n=100'
- Then use R to compute the sum of 1 through 100 using the formula $ n(n+1)/2$. What is the sum?

`@hint`
- Define `n` to be 100 in one line and simply type the formula using R code in the second line.
- Remember that in R you multiply using `*`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Here is how you compute the sum for the first 20 integers
20*(20+1)/2 

# However, we can define a variable to use the formula for other values of n
n <- 20
n*(n+1)/2

n <- 25
n*(n+1)/2

# Below, write code to calculate the sum of the first 100 integers

```

`@solution`
```{r}
# Here is how you compute the sum for the first 20 integers
20*(20+1)/2 

# However, we can define a variable to use the formula for other values of n
n <- 20
n*(n+1)/2

n <- 25
n*(n+1)/2

# Below, write code to calculate the sum of the first 100 integers 
n <- 100
n*(n+1)/2 
```

`@sct`
```{r}
test_error()
test_object("n", incorrect_msg = "Make sure that you use `n` as your variable name and that you have assigned the correct value to `n`.")
test_output_contains("5050", incorrect_msg = "You are not providing a formula that gives the correct answer. Look at the example code.")
test_student_typed("n*(n+1)/2", times = 3, not_typed_msg = "Make sure you type the equation. Look at the example code. And don't erase the example code.")
success_msg("Good job! Let`s apply this to another question.")
```
---
## Using variables 2

```yaml
type: NormalExercise
key: 213f198401
lang: r
xp: 100
skills: 1
```
What is the sum of the first 1000 positive integers? 

We can use the formula $ n(n+1)/2 $ to quickly compute this quantity.

`@instructions`
- Use the same formula as the last exercise but substitute `n`
- Instead of typing the result, use the formula and defined variable.


`@hint`
Use the same R code as you used in the first question after changing the value of `n`.

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# Below, write code to calculate the sum of the first 1000 integers 

```

`@solution`
```{r}
# Below, write code to calculate the sum of the first 1000 integers 
n <- 1000
n*(n+1)/2
```

`@sct`
```{r}
test_object("n", incorrect_msg = "Make sure that you use `n` as your variable name and that you have assigned the correct value to `n`.")
test_student_typed("n*(n+1)/2", times = 1, not_typed_msg = "Make sure you define the equation.")
success_msg("Good job! Let`s get to work on another question!")
```


---
## Functions

```yaml
type: MultipleChoiceExercise
key: 0e2a0ce37a
lang: r
xp: 100
skills: 1
```
Run the following code in the R console:

```{r, eval=FALSE}
n <- 1000
x <- seq(1,n)
sum(x)
```

Based on the result, what do you think the functions `seq` and `sum` do? You can use the help system.

`@instructions`
- sum creates a list of numbers and seq adds them up. 
- seq creates a list of numbers and sum adds them up. 
- seq computes the difference between two arguments and sum computes the sum of 1 through 1000. 
- sum always returns the same number

`@hint`
Go to R console and type `seq(1,5)`. See what you get. Then got the R console again and type `sum(seq(1,5))`. Change the `5` to other numbers.
You can also type `?` before a function to get help.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try again! Read the choices carefully. Try again."
msg2 = "Well done. Proceed to the next exercise."
msg3 = "There's no specification of 1 to 1000 for sum. Try again."
msg4 = "Read the help file for sum by typing `?sum` in the R console."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Nested function calls 1

```yaml
type: NormalExercise
key: 7df9567ad1
lang: r
xp: 100
skills: 1
```
In math and programming we say we evaluate a function when we replace arguments with specific values. So if we type `log2(16)` we evaluate the `log2` function to get the log to the base 2 of `16` which is `4`. 

In R it is often useful to evaluate a function inside another function. 
For example, `sqrt(log2(16))` will calculate the log to the base 2 of 16 and then compute the square root of that value. 
So the first evaluation gives a 4 and this gets evaluated by `sqrt` to give the final answer of 2.  


`@instructions`
- Use one line of code to compute the log, to the base 10, of the square root of 100. 
- Make sure your code includes the `log10` and `sqrt` functions.


`@hint`
The `log10` function call should be the argument for the `sqrt` function. 

`@pre_exercise_code`
```{r}
# no pec
```

`@sample_code`
```{r}
# log to the base 2 
log2(16)

# sqrt of the log to the base 2 of 16:
sqrt(log2(16))

# Compute log to the base 10 (log10) of the sqrt of 100. Do not use variables.
```

`@solution`
```{r}
# log to the base 2 
log2(16)

# sqrt of the log to the base 2 of 16:
sqrt(log2(16))

# Compute log to the base 10 (log10) of the sqrt of 100. Do not use variables.
log10(sqrt(100))
```

`@sct`
```{r}
test_error() 
test_student_typed (c("log10(sqrt(100))", "log(sqrt(100), base=10)"), times = 1, not_typed_msg = "Make sure you use the right base for the log and put the sqrt function in parenthesis. Remember not to use variables.")
success_msg("Very good!")
```


---
## Nested functions call 2

```yaml
type: MultipleChoiceExercise
key: b755d1a929
lang: r
xp: 50
skills: 1
```
Which of the following will always return the numeric value stored in `x`? You can try out examples and use the help system in the R console.

`@instructions`
- `log(10^x)`
- `log10(x^10)`
- `log(exp(x))`
- `exp(log(x, base = 2))`

`@hint`
You are looking for the case in which a function and its inverse are called sequentially. You can also try out each one on the console: assign a numeric value to `x` then try out each of the lines of code given in the choices.

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Try again! Note that `log(10^1)` is not 1 but the natural log of 10."
msg2 = "Try again! Note that `log(1^10)` is 0 not 1."
msg3 = "Well done. Proceed to the next exercise"
msg4 = "Try again! Note that `exp(log(2, base = 2)` is 0 not 2."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## End of Assessment 1

```yaml
type: PlainMultipleChoiceExercise
key: 5d1f828748
lang: r
xp: 0
skills: 1
```
This is the end of the programming assignment for this section. 

You can now close this window to go back to the <a href='https://courses.edx.org/courses/course-v1:HarvardX+PH125.1x+2T2017/courseware/cfded5c208bc4e379606cb712cc54f25/5ba06674d0be41b99185b947e09e889b/?child=first'>course</a>.

If you want to continue the assessments without watching the videos, you can click on the arrow above to get the next exercise or hit Ctrl-K. WARNING: if you continue the assessments by clicking on the arrow to get the next exercise or by hitting Ctrl-K your assessments may NOT get scored.

`@instructions`
- Awesome
- Nope

`@hint`
- No hint necessary!

`@pre_exercise_code`
```{r}
# no pec
```

`@sct`
```{r}
msg1 = "Great! Now navigate back to the course on edX!"
msg2 = "Now navigate back to the course on edX!"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2))
```
